Gender Dynamics in Financial Decision-Making: Unveiling the Money Illusion Effect in Brazil

### 2024-02-15 outputs/paper/paper.qmd
final check

Our findings reveal distinct gender differences in economic decision-making across various scenarios related to money illusion in the Brazilian context. Both men and women demonstrated a consensus on economic terms, with a preference for "Carolina," indicating a shared focus on financial stability. However, differences emerged in job attractiveness and seller preferences, suggesting gender-specific strategies and priorities in economic exchanges. Women exhibited a cautious approach to purchasing and a proactive selling strategy, reflecting a risk-averse attitude, while men showed a balanced approach to buying and selling decisions. In terms of contract preferences, both genders favored risk-free contracts in real terms, but men displayed a greater willingness to take risks under nominal terms. These findings highlight the multifaceted nature of economic decision-making, where financial stability intersects with career aspirations and risk tolerance in gender-specific ways.

@@

The data were collected via an online survey platform of and include a range of variables:

@@

This study serves as a reproduction and further exploration of the seminal work on the money illusion effect, particularly examining the impact of gender on financial decision-making. Utilizing identical datasets as the original paper, our analysis narrows the focus to the variable of gender.

@@

To contextualize the study within the Brazilian socio-economic environment, Brazilian researchers tailored the survey instrument while maintaining the core essence of the original study. A cross-sectional online survey was conducted with the prerequisites that participants be at least 18 years old and give their consent. The study received ethical clearance from the Pontifical Catholic University of São Paulo's Research Ethics Committee in 2021.

>>>>

Our findings reveal distinct gender differences in economic decision-making across various scenarios related to money illusion in the Brazilian context. Both men and women demonstrated a consensus on economic terms, with a preference for "Maria," indicating a shared focus on financial stability. However, differences emerged in job attractiveness and seller preferences, suggesting gender-specific strategies and priorities in economic exchanges. Women exhibited a cautious approach to purchasing and a proactive selling strategy, reflecting a risk-averse attitude, while men showed a balanced approach to buying and selling decisions. In terms of contract preferences, both genders favored risk-free contracts in real terms, but men displayed a greater willingness to take risks under nominal terms. These findings highlight the multifaceted nature of economic decision-making, where financial stability intersects with career aspirations and risk tolerance in gender-specific ways.

@@

The data were collected via an online survey platform and include a range of variables that capture socio-demographic information and economic decision-making scenarios. These scenarios are designed to assess susceptibility to the money illusion, a cognitive bias where individuals assess the value of money based on its nominal value, overlooking the impact of inflation. This dataset is particularly relevant for understanding how this bias manifests in a Brazilian cultural setting and how it interacts with gender differences in economic decision-making.

@@

This study serves as a reproduction and further exploration of the seminal work on the money illusion effect, particularly examining the impact of gender on financial decision-making. Utilizing identical datasets as the original paper, our analysis narrows the focus to the variable of gender. For a sample of this dataset, please refer to the appendix, which includes data tables for each research question.

@@

To contextualize the study within the Brazilian socio-economic environment, Brazilian researchers tailored the survey instrument while maintaining the core essence of the original study. A cross-sectional online survey was conducted with the prerequisites that participants be at least 18 years old and give their consent. The study received ethical clearance from the Pontifical Catholic University of São Paulo's Research Ethics Committee in 2021 [@Santiago_Feldman_Ferreira_Batistuzzo_Fatori_Seda_Bastos_2023].
